---
title: "Solutions W1 & W2"
author: "Juan Rivillas"
date: "09/09/2022"
output:
  pdf_document: default
  html_document: default
---

#### This R Markdown contains potential soluctions (there are manyof them you may apply) to the exercise of the Weeks 1 and 2: Exploratory data analsysis and correlations. 

Note: This exercise is for your training and it has not grade.

Visualise the distribution of the new variable `Age`.

```{r}
gghistostats(
  data       = ineqdata,
  x          = age,
  title      = "Histogram with distribution and mean Age of population",
  test.value = 12,
  binwidth   = 6
)

```

Find the mestizo population in the data, then identify females with diabetes

```{r}

table_gender <- table1(~ ethnic_group + diabetes | gender, data=ineqdata)
table_gender 
```

### Exploring structure of data

We can look and extract information about characteristics of data using str() and skim ()

```{r}
str(ineqdata)
skim (ineqdata)
```


### Table1 ()

Use `table1()` to generate a table with socioeconomic characteristics of population filtered by gender

```{r}

table_sep <- table1(~ education + childhood_poverty + malnutrition + migration + extreme_weather + inequality + political_party | gender, data=ineqdata)
table_sep

```

If you want to save this table in your machine you may used code as follows:

```{r}
export2word( x = ineq2, file = '/Users/macbookpro/Documents/PhD Project/Data/SABE/Outputs/ineq2.docx')
#change path file according your working directory in your machine
```


## gghistostats ()

Use gghistostats () function to look at the distribution of HD (homeostatic Dysregulation) and gender of participants plotting a histogram.


```{r}
ggbetweenstats(
  data  = ineqdata,
  x     = gender,
  y     = hd, fill = gender,
  title = "Distribution of HD (homeostatic Dysregulation) and gender of participants"
)

```

## ggbetweenstats ()

Find the mean of adult SEP (inequality) separately females and males using a violing plot

```{r}

ggbetweenstats(
  data  = ineqdata,
  x     = gender,
  y     = inequality, fill = gender,
  title = "Distribution and mean adult inequality separately females and males"
)

```

Rename levels of diabetes : healthy unhealthy

```{r}
#Rename values in the health outcomes columns and create a vectors of diseases

ineqdata$diabetes<- factor(ineqdata$diabetes,
                levels = c("yes","no"),
                labels = c("unhealthy","healthy"))
```


```{r}
Desc(ineqdata$diabetes)
Freq(ineqdata$diabetes)  
```


```

###FIN###