---
title: "Exercise and solution Linear models"
author: "Juan Rivillas"
date: "29/09/2022"
output: html_document
---

#######################################################
###        Exercise linear regression models        ###
#######################################################


Steps

Open file: regression_data hosted in GitHub

1) Build one simple linear regression model: 
   - exposure: childhood poverty and outcome: biological age (in years).
   - control by gender
    
2) Model Fitting Q-Q plot and all four default diagnostic plots. 
    Search for Best Models
  - Check the scatter and box plots
  - Check MSE, Rsquared and AIC

3) Evaluating the quality of the Linear Model.

4) Prepare your report and intepretate outcomes
  - Plot simple linear regression model
  - Tabulate linear regression model
  
  
**Solution exercise regression modelling**
  
1) Build one simple linear regression model:   
```{r}
#Building a simple linear regression model imposed on the data
lm.exe <-lm(formula = biological_age  ~ low_childhood_sep, data = regression_data)

#check coefficient model
lm.exe 
```

Linear model of low childhood sep controlling by gender
```{r}
lm.exe.gen <- lm(biological_age ~ low_childhood_sep + gender, data=regression_data)
lm.exe.gen
```


2) Model Fitting Q-Q plot and all four default diagnostic plots. 
```{r}
#The Q-Q plot for the LR model developed using the Int2000 data
qqnorm(resid(lm.exe)) > qqline(resid(lm.exe))
```
  
```{r}
#Figure: All four default diagnostic plots for the SLR model developed using the Data
par(mfrow=c(2,2)) > plot(lm.exe)
```


3) Evaluating the quality of the Linear Model displaying information with summary()

```{r}
summary (lm.exe.gen)
```

4) Prepare your report and intepretate outcomes

```{r}
#Tabulate simple regression model and extract results using tab_model().
tab_model(lm.exe.gen)
```

```{r}
#Plot coefficients and p values using library(ggstatsplot) (My favorite) #
ggcoefstats(lm.exe.gen)
```


###FIN####